\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}

\title{Global Warming: Causes, Evidence, and Effects}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Introduction}
\textit{"Your wallet and lifestyle will be affected by what people believe about global warming."} - Professor

\section*{How Does Global Warming Occur?}
\begin{itemize}
    \item Sunlight, part of the electromagnetic spectrum, passes through the atmosphere.
    \item The Earth's surface absorbs sunlight and emits it as infrared radiation.
    \item Greenhouse gases absorb infrared radiation and re-emit it, warming the surface.
    \item This process, called the \textbf{greenhouse effect}, traps heat and raises temperatures.
\end{itemize}

\section*{Evidence for Global Warming}
\subsection*{CO\textsubscript{2} Levels in the Atmosphere}
\begin{itemize}
    \item Ice cores reveal atmospheric composition over the past 800,000 years.
    \item Historical CO\textsubscript{2} levels ranged between 180 and 280 ppm.
    \item Current CO\textsubscript{2} level: about 400 ppm.
\end{itemize}

\subsection*{Ice Core Data}
\begin{itemize}
    \item Air bubbles in ice provide records of past atmospheric conditions.
    \item Ratios of Oxygen-18 to Oxygen-16 indicate historical temperatures:
    \begin{itemize}
        \item Higher ratios suggest warmer periods.
        \item Lower ratios suggest cooler periods.
    \end{itemize}
\end{itemize}

\subsection*{Carbon Isotope Analysis}
\begin{itemize}
    \item Carbon isotopes reveal shifts in the atmosphere since the Industrial Revolution:
    \begin{itemize}
        \item Fossil fuels, depleted of Carbon-14, have altered isotope ratios.
    \end{itemize}
    \item Volcanic emissions are not the main source of rising CO\textsubscript{2}.
\end{itemize}

\section*{Effects of Global Warming}
\begin{itemize}
    \item \textbf{Extreme Weather}: More frequent and intense storms.
    \item \textbf{Disasters}:
    \begin{itemize}
        \item Greater property damage.
        \item Higher mortality rates.
        \item Worsening droughts, hunger, and migration crises.
    \end{itemize}
    \item In 2022, 37.15 billion tons of CO\textsubscript{2} were released globally.
\end{itemize}

\section*{Proposed Solutions}
\begin{itemize}
    \item Reduce CO\textsubscript{2} emissions to zero.
    \item Balance mitigation with economic feasibility:
    \begin{itemize}
        \item \textit{"The fence should be as high as it needs to be."} - Professor
    \end{itemize}
\end{itemize}

\section*{Challenges to Implementation}
\begin{itemize}
    \item Uncertainty in natural CO\textsubscript{2} absorption rates.
    \item Political disagreements:
    \begin{itemize}
        \item \textit{"Focus on science, not politics."} - Professor
    \end{itemize}
\end{itemize}

\section*{Professor's Reflections}
\begin{itemize}
    \item Correlation between CO\textsubscript{2} and industrial activity is not direct proof.
    \item Conclusions must be based on rigorous testing, not patterns alone.
    \item Anecdotes:
    \begin{itemize}
        \item \textit{"Guidance on masks changed over time."} - Professor
    \end{itemize}
\end{itemize}

\end{document}
